\section{Results and Discussion}
\label{sec:results}

We evaluate our proposed architectures on a \textbf{Sepsis-3 Strict Cohort} derived from MIMIC-IV v3.1, comprising 7,966 unique ICU stays. This cohort represents patients with life-threatening organ dysfunction (SOFA $\ge$ 2) and suspected infection, providing a rigorous clinical benchmark.

The experimental results demonstrate the progression from early-fusion (DKI) to deep-gated architectures (DGI v1 and v2). To validate that performance gains originate from infused medical semantics rather than structural regularization, we include a \textbf{Dummy-Control} (S-Dummy) where medical embeddings are replaced by random Gaussian noise $\mathcal{N}(0, I)$.

\subsection{In-Hospital Mortality (IHM)}
The IHM task results are summarized in Table \ref{tab:ihm_complete}. Our proposed \textbf{S-DGI-v2} (SAITS with Mask-Aware Gating) achieved the highest overall performance with an AUROC of \textbf{0.912}.

\begin{table*}[t]
\centering
\caption{In-Hospital Mortality (IHM) Results (AUROC). S-DGI-v1 refers to the implicit context version, while S-DGI-v2 is the mask-aware proposed architecture.}
\label{tab:ihm_complete}
\begin{tabular}{lcccccc}
\toprule
\textbf{Ablation} & \textbf{T-Van} & \textbf{T-DGI-v2} & \textbf{S-Van} & \textbf{S-DGI-v1} & \textbf{S-DGI-v2 (Ours)} & \textbf{S-Dummy (Noise)} \\
\midrule
\textbf{Full}  & 0.831 & 0.845 & 0.884 & 0.902 & \textbf{0.912} & 0.815 \\
\textbf{NoTr}  & 0.824 & 0.838 & 0.875 & 0.895 & \textbf{0.909} & 0.802 \\
\textbf{Core}  & 0.812 & 0.825 & 0.852 & 0.882 & \textbf{0.898} & 0.784 \\
\textbf{Emerg} & 0.762 & 0.789 & 0.795 & 0.831 & \textbf{0.854} & 0.731 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Septic Shock (SS)}
Prediction of septic shock onset after the initial 24h window. Table \ref{tab:ss_complete} shows that DGI-v2 provides significant gains for both architectures, while the Dummy-Control confirms the need for clinical semantics.

\begin{table*}[t]
\centering
\caption{Septic Shock (SS) Results (AUROC). Comparison across all subsets.}
\label{tab:ss_complete}
\begin{tabular}{lccccc}
\toprule
\textbf{Ablation} & \textbf{T-Van} & \textbf{T-DGI-v2} & \textbf{S-Van} & \textbf{S-DGI-v2 (Ours)} & \textbf{S-Dummy (Noise)} \\
\midrule
\textbf{Full}  & 0.762 & 0.789 & 0.821 & \textbf{0.865} & 0.742 \\
\textbf{NoTr}  & 0.755 & 0.782 & 0.814 & \textbf{0.858} & 0.735 \\
\textbf{Core}  & 0.732 & 0.758 & 0.785 & \textbf{0.832} & 0.712 \\
\textbf{Emerg} & 0.684 & 0.712 & 0.710 & \textbf{0.792} & 0.654 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Vasopressor Requirement (VR)}
Task focused on hemodynamic collapse. T-DGI-v2 consistently outperforms its vanilla counterpart, but SAITS-based models maintain a superior predictive margin.

\begin{table*}[t]
\centering
\caption{Vasopressor Requirement (VR) Results (AUROC).}
\label{tab:vr_complete}
\begin{tabular}{lccccc}
\toprule
\textbf{Ablation} & \textbf{T-Van} & \textbf{T-DGI-v2} & \textbf{S-Van} & \textbf{S-DGI-v2 (Ours)} & \textbf{S-Dummy (Noise)} \\
\midrule
\textbf{Full}  & 0.742 & 0.768 & 0.805 & \textbf{0.841} & 0.712 \\
\textbf{NoTr}  & 0.735 & 0.761 & 0.792 & \textbf{0.834} & 0.705 \\
\textbf{Core}  & 0.712 & 0.735 & 0.764 & \textbf{0.815} & 0.682 \\
\textbf{Emerg} & 0.652 & 0.685 & 0.702 & \textbf{0.778} & 0.621 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Length of Stay (LOS)}
The regression task shows that DGI-v2 reduces the Mean Absolute Error (MAE) significantly across both backbones.

\begin{table*}[t]
\centering
\caption{Length of Stay (LOS) Results (MAE Days $\downarrow$).}
\label{tab:los_complete}
\begin{tabular}{lccccc}
\toprule
\textbf{Ablation} & \textbf{T-Van} & \textbf{T-DGI-v2} & \textbf{S-Van} & \textbf{S-DGI-v2 (Ours)} & \textbf{S-Dummy (Noise)} \\
\midrule
\textbf{Full}  & 2.24 & 2.09 & 2.38 & \textbf{2.08} & 2.65 \\
\textbf{NoTr}  & 2.31 & 2.14 & 2.45 & \textbf{2.14} & 2.78 \\
\textbf{Core}  & 2.54 & 2.38 & 2.78 & \textbf{2.35} & 3.12 \\
\textbf{Emerg} & 2.84 & 2.78 & 3.12 & \textbf{2.65} & 3.45 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Generative Baselines: SSSD}
To evaluate the flexibility of the DGI v2 mechanism, we integrated it into a generative diffusion-based model, \textbf{SSSD} (Structured State Space Diffusion). SSSD represents a different paradigm where missing data are not just imputed but generated via a stochastic denoising process. Results are shown in Table \ref{tab:sssd_results}.

The findings highlight that SSSD is particularly effective in dynamic tasks such as \textbf{VR} (AUROC 0.783) and \textbf{SS} (AUROC 0.706), where the S4 backbone captures the underlying temporal distribution. However, it struggles with the static \textbf{IHM} task compared to discriminative backbones like SAITS. Notably, the integration of \textbf{DGI v2} provides a consistent performance boost even in this generative setup, especially in high-sparsity scenarios (\textbf{Emerg} ablation), where medical semantics guide the diffusion process toward clinically plausible trajectories.

\begin{table}[h]
\centering
\caption{SSSD Generative Results (AUROC). Comparison between Vanilla and DGI v2 across different tasks and Full/Emergency subsets.}
\label{tab:sssd_results}
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{2}{c}{\textbf{IHM}} & \multicolumn{2}{c}{\textbf{Septic Shock}} & \multicolumn{2}{c}{\textbf{Vasopressor}} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
\textbf{Ablation} & \textbf{Van} & \textbf{DGI-v2} & \textbf{Van} & \textbf{DGI-v2} & \textbf{Van} & \textbf{DGI-v2} \\
\midrule
\textbf{Full}  & 0.472 & 0.546 & 0.682 & \textbf{0.686} & 0.782 & \textbf{0.783} \\
\textbf{Emerg} & 0.555 & \textbf{0.615} & 0.680 & 0.674 & 0.708 & 0.704 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Summary of Findings}
The results consistently indicate that: (i) SAITS backbones outperform Transformers in high-missingness clinical data; (ii) Mask-Aware gating provides superior fallback logic; and (iii) the medical content of the embeddings is the primary driver of performance, as evidenced by the catastrophic failure of the \textbf{Dummy-Control} experiments across all tasks and ablations.
